### 2018-01-27 includes/problems.tex
Adding more sections. Adding references.

    We generated several suites of string and regex problems using \fuzzer{}. \footnote{For review anonymity, we removed the link to the problem repository. It will be provided separately in the artifact submission track and will appear in the final version of the paper.} Below, we describe the \texttt{generated} suite that we generated using \generator{}: \\

    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}    & \textbf{Problems have ...} \\ \hline
        Concats          & A single, deep tree of concats. \\ \hline
        Concats-Extracts & A single concat tree, and random character extractions. \\ \hline
        Lengths          & A single, large length constraint on a variable. \\ \hline
        Lengths-Concats  & Variables and their concats, each with fixed lengths. \\ \hline
        Overlaps         & A formula in the form A.X = X.B. \\ \hline
        Regex            & A test for regex membership of arbitrary complexity. \\ \hline
        Many-Regexes     & A test for membership in many random regexes. \\ \hline
        Regex-Deep       & A test for membership in a regex with many nested operators. \\ \hline
        Regex-Pair       & A test for membership in one regex, but not another. \\ \hline
        Regex-Lengths    & A test for membership in a regex, and a length constraint. \\ \hline
    \end{tabular}
    \hfill \\

    \todo{Federico: fill in descriptions. First draft - Federico}

    The \texttt{transformed} suite contains benchmarks derived from existing seed problems by iteratively applying \transformer{}.
    Every problem in the suite is named according to its seed and the transformations it undertook. 
    For example, a hypothetical file named ``seed-fuzz-graft.smt2'' began as a file named ``seed.smt'' and was transformed twice:
    first by fuzz, and then by graft.

    The suite is divided into two categories. For an overview see Table~\ref{tbl:transformed}.

    \begin{table}
    \begin{tabular}{|l|l|r|}
        \hline
        \textbf{Name} & \textbf{Seed}                            & \textbf{Quantity} \\ \hline
        Amazon        & Two industrial regex membership problems & 516\\ \hline
        Regex         & Z3str2 regular expression test suite     & 10,577\\ \hline
    \end{tabular}
    \caption{\transformer{} generated problem sets.}
    \label{tbl:transformed}
    \end{table}

@@

        The first bug we found caused \us{} to perform poorly when the equivalence class of arithmetic terms is very large (e.g. where many strings are equal to the empty string, and thus have length 0). The slowdown was caused by a loop over the equivalence class of these terms to search for an integer constant. The fix was to check the root of the equivalence class for the integer constant, as Z3 makes "interpreted terms" (i.e. constants) the root of equivalence classes whenever possible. This bug was last observed in commit 6308636, and was fixed in commit 3865c45.

        The second bug we found caused \us{} to perform poorly when a variable had a large length constraint. The slowdown was caused by the solver performing a linear search for a satisfying length when producing a model for the variable. This bug was an inconsistency between the implementation and the behaviour described in the \us{} paper \cite{z3str3}. This bug was last observed in commit 66bc68f, and was fixed in commit 7b536e9.

>>>>

    We generated several suites of string and regex problems using \fuzzer{}. \footnote{For review anonymity, we removed the link to the problem repository. It will be provided separately in the artifact submission track and will appear in the final version of the paper.} Table~\ref{tbl:generated} describes the \texttt{generated} suite that we generated using \generator{}.

    \begin{table}
        \centering
        \begin{tabular}{|l|l|}
            \hline
            \textbf{Name}             & \textbf{Problems have ...} \\ \hline
            \textit{Concats}          & A single, deep tree of concats. \\ \hline
            \textit{Concats-Extracts} & A single concat tree, and random character extractions. \\ \hline
            \textit{Lengths}          & A single, large length constraint on a variable. \\ \hline
            \textit{Lengths-Concats}  & Variables and their concats, each with fixed lengths. \\ \hline
            \textit{Overlaps}         & A formula of the form A.X = X.B. \\ \hline
            \textit{Regex}            & A test for regex membership of arbitrary complexity. \\ \hline
            \textit{Many-Regexes}     & A test for membership in many random regexes. \\ \hline
            \textit{Regex-Deep}       & A test for membership in a regex with many nested operators. \\ \hline
            \textit{Regex-Pair}       & A test for membership in one regex, but not another. \\ \hline
            \textit{Regex-Lengths}    & A test for membership in a regex, and a length constraint. \\ \hline
        \end{tabular}
        \caption{\generator{}-generated problem sets.}
        \label{tbl:generated}
    \end{table}

    The \texttt{transformed} suite contains benchmarks derived from existing seed problems by iteratively applying \transformer{}. Every problem in the suite is named according to its seed and the transformations it undertook. For example, a hypothetical file named ``seed-fuzz-graft.smt2'' began as a file named ``seed.smt'' and was transformed twice: first by fuzz, and then by graft. The suite is divided into two categories. For an overview see Table~\ref{tbl:transformed}.

    \begin{table}
        \centering
        \begin{tabular}{|l|l|r|}
            \hline
            \textbf{Name}   & \textbf{Seed}                            & \textbf{Quantity} \\ \hline
            \textit{Amazon} & Two industrial regex membership problems & 516\\ \hline
            \textit{Regex}  & Z3str2 regular expression test suite     & 10,577\\ \hline
        \end{tabular}
        \caption{\transformer{}-generated problem sets.}
        \label{tbl:transformed}
    \end{table}

@@

        The first bug we found caused \us{} to perform poorly when equivalence classes of arithmetic terms was very large (e.g. where many strings were equal to the empty string, and thus had length 0). We discovered this bug using the \textit{Concats} problem suite. The slowdown was caused by a loop over the equivalence class of these terms to search for an integer constant. The fix was to check the root of the equivalence class for the integer constant, as Z3 makes "interpreted terms" (i.e. constants) the root of equivalence classes whenever possible. This bug was last observed in commit 6308636, and was fixed in commit 3865c45.

        The second bug we found caused \us{} to perform poorly when a variable had a large length constraint. We discovered this bug using the \textit{Lengths} problem suite. The slowdown was caused by the solver performing a linear search for a satisfying length when producing a model for the variable. This bug was an inconsistency between the implementation and the behaviour described in the \us{} paper \cite{z3str3}. This bug was last observed in commit 66bc68f, and was fixed in commit 7b536e9.

### 2018-01-27 includes/stringfuzz.tex
Adding more sections. Adding references.

        \item[\generator{}] \hfill \\
            This tool generates \smt{} problems. It supports several generators and options that specify its output. The generators it implements are: \\

            \begin{tabular}{|l|l|}
                \hline
                \textbf{Name} & \textbf{Generated Problems have ...} \\ \hline
                concats       & Long concats and optional random extracts. \\ \hline
                lengths       & Many variables (and their concats) with length constraints. \\ \hline
                overlaps      & An expression of the form A.X = X.B. \\ \hline
                equality      & An equality among concats, each with variables or constants. \\ \hline
                regex         & Regexes of varying complexity. \\ \hline
                random-text   & Random, likely syntactically \textit{in}valid text. \\ \hline
                random-ast    & Random, but semantically \textit{valid} text. \\ \hline
            \end{tabular}

            \hfill \\

        \item[\transformer{}] \hfill \\

            This tool transforms \smt{} problems.. It supports several transformers and options that specify its output and input. The transformers it implements are: \\

            \todo{Dmitry: maybe add notes about SAT preservation?}

            \begin{tabular}{|l|l|}
                \hline
                \textbf{Name} & \textbf{The transformer ...} \\ \hline
                fuzz          & Randomly replaces literals and operators with similar ones. \\ \hline
                graft         & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
                multiply      & Multiplies each number by N, and repeats each character N times. \\ \hline
                nop           & Does nothing. \\ \hline
                reverse       & Reverses all string literals and concat arguments. \\ \hline
                rotate        & Performs a random syntax tree rotation on two concats. \\ \hline
                translate     & Maps every character X to a different character Y. \\ \hline
                unprintable   & Randomly replaces all characters in literals with unprintable ones. \\ \hline
            \end{tabular}

            \hfill \\

        \item[\texttt{stringstats}] \hfill \\
            This tool takes as input an \smt{} problem, and outputs its properties, such as: numbers of string variables and literals, max/median syntactic depth of expressions, and max/median length of literals.

            \hfill

@@

        \fuzzer{} can generate and transform problems with regular expression constraints. The command \texttt{stringfuzzg regex} invokes the regex problem generator, and has several options to control its output. It generates a problem of the form\footnote{Variable declarations omitted for brevity.}:

        \begin{align*}
            \langle problem \rangle  & ::= \langle membership \rangle \; \langle membership \rangle * \\
            \langle membership \rangle & ::= \texttt{(assert (str.in.re X}\; R\; \texttt{))}
        \end{align*}

        where $R \in RegEx$. More simply, the problem is a set of one or more regex constraints on a single variable.

        \todo{Dmitry: further describe the properties of $R$.}

@@

        \fuzzer{} can be used to generate static benchmarks as described in Section~\ref{sec:problems}. It can also be used in a randomised fashion for exploratory debugging. As a motivating example, the following usage can feed random problems to \cvc{} until it returns an error:

        {\scriptsize\begin{verbatim}
while stringfuzzg -r random --num-terms 1000 | tee problem.smt25 | cvc4 --lang smt2.5 --tlimit=5000; do
    sleep 0
done\end{verbatim}}

>>>>

        \item[\generator{}] \hfill \\
            This tool generates \smt{} problems. It supports several generators and options that specify its output. Table~\ref{tbl:generators} describes them.
            \hfill \\

            \begin{table}
                \centering
                \begin{tabular}{|l|l|}
                    \hline
                    \textbf{Name}        & \textbf{Generated Problems have ...} \\ \hline
                    \textit{concats}     & Long concats and optional random extracts. \\ \hline
                    \textit{lengths}     & Many variables (and their concats) with length constraints. \\ \hline
                    \textit{overlaps}    & An expression of the form A.X = X.B. \\ \hline
                    \textit{equality}    & An equality among concats, each with variables or constants. \\ \hline
                    \textit{regex}       & Regexes of varying complexity. \\ \hline
                    \textit{random-text} & Random, likely syntactically \textit{in}valid text. \\ \hline
                    \textit{random-ast}  & Random, but semantically \textit{valid} text. \\ \hline
                \end{tabular}
                \caption{\generator{} generators}
                \label{tbl:generators}
            \end{table}

        \item[\transformer{}] \hfill \\
            This tool transforms \smt{} problems. It supports several transformers and options that specify its output and input. Table~\ref{tbl:transformers} describes them. The following transformers also preserve problem satisfiability: \textit{translate}, \textit{reverse}. \cite{ifaz}
            \hfill \\

            \begin{table}
                \centering
                \begin{tabular}{|l|l|}
                    \hline
                    \textbf{Name}        & \textbf{The transformer ...} \\ \hline
                    \textit{fuzz}        & Randomly replaces literals and operators with similar ones. \\ \hline
                    \textit{graft}       & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
                    \textit{multiply}    & Multiplies each number by N, and repeats each character N times. \\ \hline
                    \textit{nop}         & Does nothing. \\ \hline
                    \textit{reverse}     & Reverses all string literals and concat arguments. \\ \hline
                    \textit{rotate}      & Performs a random syntax tree rotation on two concats. \\ \hline
                    \textit{translate}   & Maps every character X to a different character Y. \\ \hline
                    \textit{unprintable} & Randomly replaces all characters in literals with unprintable ones. \\ \hline
                \end{tabular}
                \caption{\transformer{} transformers}
                \label{tbl:transformers}
            \end{table}

        \item[\texttt{stringstats}] \hfill \\
            This tool takes as input an \smt{} problem, and outputs its properties, such as: numbers of string variables and literals, max/median syntactic depth of expressions, and max/median length of literals.
            \hfill \\

@@

        \fuzzer{} can generate and transform problems with regular expression constraints. The command \texttt{stringfuzzg regex} invokes the regex problem generator, and has several options to control its output. It generates a problem of the form\footnote{Variable declarations omitted for brevity.}:
        \begin{align*}
            & \texttt{(assert (str.in.re X}\; R_0\; \texttt{))} \\
            & \texttt{(assert (str.in.re X}\; R_n\; \texttt{))}* \\
            & \texttt{(assert (<= Min (str.len X)))}? \\
            & \texttt{(assert (<= (str.len X)) Max)}?
        \end{align*}

        where $R_i \in RegEx$, and $Min, Max \in Int$. More simply, the problem is a set of one or more regex constraints on a single variable, with optional maximum and minimum lengths constraints. The regex constraints $R$ are each of the form:

        \begin{align*}
            & \texttt{(re.++}\; T_0\; T_1\; \texttt{...}\; T_n\; \texttt{)}
        \end{align*}

        and each $T_i$ is a recursive term of the form:

        \begin{align*}
            & \texttt{(re.*}\; T_{i_j}\; \texttt{) | (re.+}\; T_{i_j}\; \texttt{) | (re.union}\; T_{i_{j_1}}\; T_{i_{j_2}}\; \texttt{)}
        \end{align*}

        where $j$ a specified depth of recursion. Terms at depth 0 are regex constants. Informally, this form describes a concatenation of regex terms. Each term is a random nested regex operator (chosen from regex Kleene star, repetition, and union), up to a specified depth, terminating in a regex literal. Below is example regex (with spacing added for clarity) of depth 2 produced by this scheme:

        \begin{align*}
            & ((\texttt{a}|\texttt{b})|(\texttt{cc})+)\quad ((\texttt{ddd})*)+\quad ((\texttt{ee})+|(\texttt{fff})*)
        \end{align*}

        \fuzzer{} can also transform the instances it generates. All of the transformers in \transformer{} can operate on regex terms. They can multiply and permute regex literals (e.g. \textit{reverse}, \textit{multiply}, \textit{translate}), randomly replace regex operators (e.g. \textit{fuzz}), and manipulate the syntax tree of a regex problem (e.g. \textit{graft}, \textit{rotate}).

@@

        \fuzzer{} can be used to generate problems to be saved as static suites like those in Section~\ref{sec:problems}. It can also be used in a randomised fashion for exploratory debugging. As a motivating example, the following usage can feed random problems to \cvc{} until it returns an error:

        {\scriptsize\begin{verbatim}while stringfuzzg -r random --num-terms 1000 | tee problem.smt25 | cvc4 --lang smt2.5 --tlimit=5000; do
    sleep 0
done\end{verbatim}}
